Multimodal Information Extraction of Supermarket Leaflets

### 2025-03-15 report/main.tex
Cleaned up codebase.

\indent{\textbf{Deal Detection}.}
There are no direct approaches for detecting deals, including product image, price, name and more, in supermarket leaflets. However, the detection of specific elements in documents, or more generally, the detection of objects in images, is a widely researched area. When it comes to just the actual detection of these elements without understanding the semantic meaning, models such as YOLO or Faster R-CNN prove to be effective. These models are able to recognize objects in images and mark them with bounding boxes and further classify them. 
% TODO: Add more advanced methods for object/instant segmentation

@@

\subsection{Dataset}
An approach to obtaining deal data would be direct web scraping from supermarket websites, as these days they present their deals digitally in a structured way. However, this method is highly error-prone and difficult to generalize due to the distinct and frequently changing layouts of different supermarket websites. Additionally, many retailers do not provide direct, easily scrappable structured deal data. Instead, an alternative source has been identified. Websites that aggregate and upload full promotional leaflets from various German supermarkets on a weekly basis. These leaflets provide a more standardized input format and can be systematically downloaded using an automated script.
% TODO: add url source

@@

Object detection models can be broadly categorized into multi-stage (multi-shot) detectors and single-stage (single-shot) detectors. Multi-stage detectors, such as the R-CNN family (R-CNN, Fast R-CNN, Faster R-CNN, Mask R-CNN), follow a sequential process. First, they generate a set of region proposals, indicating potential object locations. These proposals are then classified in a second stage. While this method achieves high accuracy, it introduces significant computational overhead. Earlier versions, such as R-CNN, even relied on separate, non-learnable algorithms for proposal generation. Faster R-CNN improved efficiency by integrating the Region Proposal Network (RPN) directly into the model, eliminating the need for external proposal algorithms. However, even this approach remains computationally demanding. In contrast, single-stage detectors, such as YOLO (You Only Look Once), perform detection in a single forward pass, predicting object locations and classifications simultaneously. This significantly reduces inference time, making these models well-suited for real-time applications. YOLO, in particular, has established itself as one of the most widely used architectures due to its speed and competitive accuracy. Given the requirement of processing thousands of leaflet pages efficiently, a single-stage detector was chosen for deal detection, with a particular focus on YOLO-based architectures.

The YOLO model follows a grid-based detection approach, where the input image is divided into an SxS grid. Each grid cell is processed for predicting objects whose centers fall within it. Instead of relying on separate region proposal mechanisms, YOLO processes the image in a single forward pass, directly predicting bounding box coordinates, confidence scores, and class probabilities per grid cell. The bounding boxes are parameterized by their center coordinates, width, and height, while the confidence scores indicate the model’s certainty that a given bounding box contains an object. The class probabilities further specify which category the detected object belongs to. Traditional object detection frameworks like YOLO produce bounding boxes that encapsulate detected objects in rectangular shapes. However, supermarket leaflets contain deals that frequently overlap or feature irregular layouts, making bounding box-based detection less effective. Instance segmentation extends YOLO by incorporating an additional segmentation head that predicts pixel-level masks for each detected object. This approach enables the model to distinguish individual deals more precisely, reducing the risk of grouping multiple offers within a single bounding box. The segmentation mechanism operates by first detecting objects using the standard YOLO pipeline. Once a bounding box is identified, the model refines its prediction by generating a binary mask that determines which pixels within the bounding box belong to the detected object. The segmentation head shares feature representations with the object detection components, ensuring efficient training and inference. This method enhances the model’s ability to isolate individual deals even in cluttered layouts, making it particularly suited for supermarket leaflets where offers frequently overlap or appear with complex backgrounds.

\subsection{Experiments}
Evaluating the performance of instance segmentation models requires an analysis of how well detected objects align with ground-truth annotations. The primary metric used in this assessment is \emph{Mean Average Precision (mAP)}, which quantifies detection accuracy at different levels of overlap between predicted and actual objects. The core component of mAP is Intersection over Union (IoU). IoU is defined as the ratio of the intersection area between the predicted and ground-truth segmentation masks to their union. A higher IoU indicates a better match between the prediction and the actual object, while a lower IoU suggests that the detected region is misaligned or includes excessive background. In the case of bounding boxes, IoU is calculated based on the overlap of rectangular regions, whereas for instance segmentation, it is computed on a pixel level, capturing the true shape of the detected objects. To evaluate model performance, different IoU thresholds are applied. A commonly used metric is mAP@50, which considers a detection correct if the IoU between the predicted and ground-truth mask exceeds 50\%. This threshold provides insight into how well a model identifies objects in general but does not strictly enforce precision in localization. A more comprehensive metric is mAP@50-95, which averages precision across multiple IoU thresholds, ranging from 50\% to 95\% in increments of 5\%. This approach penalizes models that fail to produce accurate segmentations and is particularly useful for assessing how well the model distinguishes closely packed or irregularly shaped deals. The implemention details can be found in \secref{app:implementation_details}.

@@

The first set of experiments focused on comparing different YOLO variants trained on the dataset. Each model was trained for 20 epochs without additional data augmentation. The results, shown in \figref{fig:all_yolo_results} (a), present the mAP50 (mask) scores for YOLOv8 and YOLOv11 in their small, medium, and large configurations. The findings indicate that the smallest version of YOLOv8 achieved the highest validation mAP50, outperforming larger models in this setting. A second analysis, visualized in \figref{fig:all_yolo_results} (b), examines mAP50-95 scores for the same models, reinforcing the observation that a lightweight YOLO model is sufficient for this particular task. The lower performance of larger models suggests that the dataset’s structure, with relatively uniform layouts, does not require highly complex architectures to achieve strong results.

@@

Following this baseline evaluation, additional experiments investigated the impact of data augmentation on segmentation performance. Various augmentation strategies were applied, including translation, scaling, horizontal flipping, and color transformations. The goal was to determine whether these transformations improved the model’s ability to generalize across different leaflet layouts. Each augmented model was trained for ten additional epochs, and the results are displayed in \figref{fig:yolo8_results}. While certain augmentation techniques led to slight improvements in segmentation accuracy, particularly for models trained with random translations, the overall best-performing model remained the one trained without any augmentation. This outcome suggests that the dataset already contains sufficient variation to ensure robust generalization without the need for extensive artificial augmentations.

The final model selection was based on the combination of high segmentation accuracy and computational efficiency. Given the results, \emph{YOLOv8s} was chosen as the preferred model, as it consistently outperformed more complex variants while maintaining a lower computational cost. The model’s ability to detect deals with high precision makes it a strong candidate for further applications. And this is particularly important because everything else builds on it.

@@

\section{Application}
\label{sec:application}
The ability to detect deals from a wide range of supermarket leaflets provides the foundation for extracting structured information and building a centralized deal database. If all components, from deal detection to information extraction, function reliably, this enables the creation of a comprehensive repository containing deals from various supermarkets, based solely on the content of their weekly unstructured leaflets. Such a structured database allows for numerous applications, including price trend analysis, intelligent shopping list generation and deal comparison across stores. The immediate objective of this project was to develop a system that could automatically collect, process and present extracted deals in a structured manner. The resulting application is designed to scrape supermarket leaflets weekly, detect deals using the trained YOLO-based model and extract relevant metadata, including product names and prices. These extracted deals are then stored in a structured format within a database and made accessible through a web-based interface.

@@

    Firstly, enhancing the performance and runtime of the deal detection module is crucial. This involves optimizing the existing YOLO-based model and exploring alternative architectures that may offer better accuracy and efficiency. Also, the size of the dataset can be increased by including more leaflets from different supermarkets, which would improve the model's generalization capabilities.

@@

\bibliography{references}
\newpage

>>>>

\indent{\textbf{Deal Detection}.}
There are no direct approaches for detecting deals, including product image, price, name and more, in supermarket leaflets. However, the detection of specific elements in documents, or more generally, the detection of objects in images, is a widely researched area. When it comes to just the actual detection of these elements without understanding the semantic meaning, models such as YOLO (You Only Look Once) \cite{redmonYouOnlyLook2016} or Faster R-CNN \cite{renFasterRCNNRealTime2016} prove to be effective. These models are able to recognize objects in images and mark them with bounding boxes and further classify them. More advanced methods, such as Mask R-CNN \cite{heMaskRCNN2018} or additional segmentation heads for YOLO like proposed in \cite{mohamedINSTAYOLORealTimeInstance2024}, can even segment the objects in the image. Furthermore, more recent models, such as DETR \cite{zhuDeformableDETRDeformable2021} or RT-DETR \cite{zhaoDETRsBeatYOLOs2024a}, have shown promising results in object detection tasks using transformer-based architectures. So there is a wide range of models available for the task of object detection and segmentation in general, which can be investigated for the task of deal detection in supermarket leaflets.

@@

\subsection{Dataset}
An approach to obtaining deal data would be direct web scraping from supermarket websites, as these days they present their deals digitally in a structured way. However, this method is highly error-prone and difficult to generalize due to the distinct and frequently changing layouts of different supermarket websites. Additionally, many retailers do not provide direct, easily scrappable structured deal data. Instead, an alternative source has been identified. Websites that aggregate and upload full promotional leaflets from various German supermarkets on a weekly basis. These leaflets provide a more standardized input format and can be systematically downloaded using an automated script. For this purpose, a web scraper was developed to download the leaflets from the website \url{https://www.prospektangebote.de/}.

@@

Object detection models can be broadly categorized into multi-stage (multi-shot) detectors and single-stage (single-shot) detectors. Multi-stage detectors, such as the R-CNN family (R-CNN \cite{girshickRichFeatureHierarchies2014a}, Fast R-CNN \cite{girshickFastRCNN2015}, Faster R-CNN, Mask R-CNN), follow a sequential process. First, they generate a set of region proposals, indicating potential object locations. These proposals are then classified in a second stage. While this method achieves high accuracy, it introduces significant computational overhead. Earlier versions, such as R-CNN, even relied on separate, non-learnable algorithms for proposal generation \cite{uijlingsSelectiveSearchObject2013}. Faster R-CNN improved efficiency by integrating the Region Proposal Network (RPN) directly into the model, eliminating the need for external proposal algorithms. However, even this approach remains computationally demanding. In contrast, single-stage detectors, such as YOLO, perform detection in a single forward pass, predicting object locations and classifications simultaneously. This significantly reduces inference time, making these models well-suited for real-time applications. YOLO, in particular, has established itself as one of the most widely used architectures due to its speed and competitive accuracy. Given the requirement of processing thousands of leaflet pages efficiently, a single-stage detector was chosen for deal detection, with a particular focus on YOLO-based architectures.

The YOLO model follows a grid-based detection approach, where the input image is divided into an $S \times S$ grid. Each grid cell is processed for predicting objects whose centers fall within it. Instead of relying on separate region proposal mechanisms, YOLO processes the image in a single forward pass, directly predicting bounding box coordinates, confidence scores, and class probabilities per grid cell. The bounding boxes are parameterized by their center coordinates, width, and height, while the confidence scores indicate the model's certainty that a given bounding box contains an object. The class probabilities further specify which category the detected object belongs to. Traditional object detection frameworks like YOLO produce bounding boxes that encapsulate detected objects in rectangular shapes. However, supermarket leaflets contain deals that frequently overlap or feature irregular layouts, making bounding box-based detection less effective. Instance segmentation extends YOLO by incorporating an additional segmentation head that predicts pixel-level masks for each detected object. This approach enables the model to distinguish individual deals more precisely, reducing the risk of grouping multiple offers within a single bounding box. The segmentation mechanism operates by first detecting objects using the standard YOLO pipeline. Once a bounding box is identified, the model refines its prediction by generating a binary mask that determines which pixels within the bounding box belong to the detected object. The segmentation head shares feature representations with the object detection components, ensuring efficient training and inference. This method enhances the model's ability to isolate individual deals even in cluttered layouts, making it particularly suited for supermarket leaflets where offers frequently overlap or appear with complex backgrounds.

\subsection{Experiments}
Evaluating the performance of instance segmentation models requires an analysis of how well detected objects align with ground-truth annotations. The primary metric used in this assessment is \emph{Mean Average Precision (mAP)}, which quantifies detection accuracy at different levels of overlap between predicted and actual objects. The core component of mAP is Intersection over Union (IoU). IoU is defined as the ratio of the intersection area between the predicted and ground-truth segmentation masks to their union. A higher IoU indicates a better match between the prediction and the actual object, while a lower IoU suggests that the detected region is misaligned or includes excessive background. In the case of bounding boxes, IoU is calculated based on the overlap of rectangular regions, whereas for instance segmentation, it is computed on a pixel level, capturing the true shape of the detected objects. To evaluate model performance, different IoU thresholds are applied. A commonly used metric is mAP@50, which considers a detection correct if the IoU between the predicted and ground-truth mask exceeds 50\%. This threshold provides insight into how well a model identifies objects in general but does not strictly enforce precision in localization. A more comprehensive metric is mAP@50-95, which averages precision across multiple IoU thresholds, ranging from 50\% to 95\% in increments of 5\%. This approach penalizes models that fail to produce accurate segmentations and is particularly useful for assessing how well the model distinguishes closely packed or irregularly shaped deals.

For the experiments, the \emph{YOLOv8} and \emph{YOLOv11} models were selected, as they are the most recent and widely used architectures in the YOLO family which support instance segmentation with available pretrained weights. The implemention details can be found in \secref{app:implementation_details}.

@@

The first set of experiments focused on comparing different YOLO variants trained on the dataset. Each model was trained for 20 epochs without additional data augmentation. The results, shown in \figref{fig:all_yolo_results} (a), present the mAP50 (mask) scores for YOLOv8 and YOLOv11 in their small, medium, and large configurations. The findings indicate that the smallest version of YOLOv8 achieved the highest validation mAP50, outperforming larger models in this setting. A second analysis, visualized in \figref{fig:all_yolo_results} (b), examines mAP50-95 scores for the same models, reinforcing the observation that a lightweight YOLO model is sufficient for this particular task. The lower performance of larger models suggests that the dataset's structure, with relatively uniform layouts, does not require highly complex architectures to achieve strong results.

@@

Following this baseline evaluation, additional experiments investigated the impact of data augmentation on segmentation performance. Various augmentation strategies were applied, including translation, scaling, horizontal flipping, and color transformations. The goal was to determine whether these transformations improved the model's ability to generalize across different leaflet layouts. Each augmented model was trained for ten additional epochs, and the results are displayed in \figref{fig:yolo8_results}. While certain augmentation techniques led to slight improvements in segmentation accuracy, particularly for models trained with random translations, the overall best-performing model remained the one trained without any augmentation. This outcome suggests that the dataset already contains sufficient variation to ensure robust generalization without the need for extensive artificial augmentations.

The final model selection was based on the combination of high segmentation accuracy and computational efficiency. Given the results, YOLOv8s was chosen as the preferred model, as it consistently outperformed more complex variants while maintaining a lower computational cost. The model's ability to detect deals with high precision makes it a strong candidate for further applications. And this is particularly important because everything else builds on it.

@@

\section{Application}
\label{sec:application}
The ability to detect deals from a wide range of supermarket leaflets provides the foundation for extracting structured information and building a centralized deal database. If all components, from deal detection to information extraction, function reliably, this enables the creation of a comprehensive repository containing deals from various supermarkets, based solely on the content of their weekly unstructured leaflets. Such a structured database allows for numerous applications, including price trend analysis, intelligent shopping list generation and deal comparison across stores. The immediate objective of this project was to develop a system that could automatically collect, process and present extracted deals in a structured manner. The resulting application is designed to scrape supermarket leaflets weekly, detect deals using the trained YOLO model and extract relevant metadata, including product names and prices. These extracted deals are then stored in a structured format within a database and made accessible through a web-based interface.

@@

    Firstly, enhancing the performance and runtime of the deal detection module is crucial. This involves optimizing the existing YOLO-based model and exploring alternative architectures that may offer better accuracy and efficiency like transformer based architectures like DETR. Also, the size of the dataset can be increased by including more leaflets from different supermarkets, which would improve the model's generalization capabilities.

@@

\newpage
\bibliography{references}
\newpage
